---
title: "covid 19"
author: "Briann Brazill"
date: "4/9/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this


```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
Among the females, a greater percentage of them are alive, and among the males a greater percentage of them have died. 
  





  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The nummerical summary shows almost twice as many males died as females

## Inferential results 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The P value is 0.0009. This is a small number so we reject the null hypothesis. We can conclude that there is a relationship between sex and whether you die of covid. 


## Conclusions

We found there is relationship between sex and whether you die of covid 19 in korea. Men are more likely to die than woman. 